Toward early-warning detection of gravitational waves from compact binary coalescence

### 2014-05-06 article.tex
Make two multiline equations single-line
Adjust spacing for emulateapj

From the combination of the time slice decomposition in
Equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
Equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
up to time slice $s$ using the recurrence relation,
%
% orthogonal decomposition filtering
%
\begin{multline}
	\rho_i^s [k] =%
		% Interpolation SNR
		\overbrace{
			\left(H^\uparrow \rho_i^{s+1}\right)[k]
		}^\textrm{\clap{S/N from previous time slices}} \\
		% Plus ...
		+
		% Reconstruction
		\underbrace{
			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
		}_\textrm{\clap{reconstruction}}
		% Orthogonal FIR filter
		\overbrace{
			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
\end{multline}
%
%
Observe that the early-warning output for time slice 0, $\rho_i^0[k]$,
approximates the \SNR\ of the original templates.
%
%
\begin{figure*}[h!]
	\begin{center}
		\includegraphics{figures/lloid-diagram}
		\caption{\label{fig:pipeline} Schematic of \lloid\ pipeline illustrating
signal flow.  Circles with arrows represent interpolation
\protect\includegraphics{figures/upsample-symbol} or decimation
\protect\includegraphics{figures/downsample-symbol}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol}.  Squares
\protect\includegraphics{figures/fir-symbol} stand for \fir\ filters.  Sample
rate decreases from the top of the diagram to the bottom.  In this diagram, each
time slice contains three \fir\ filters that are linearly combined to produce
four output channels.  In a typical pipeline, the number of \fir\ filters is
much less than the number of output channels.}
	\end{center}
\end{figure*}
%
%
The signal flow diagram in Figure~\ref{fig:pipeline} illustrates this
recursion relation as a multirate filter network with a number of early-warning outputs.

@@

Taking into account the decimation of the detector data, the orthogonal \fir\
filters, the reconstruction of the time-sliced templates, the interpolation of
\SNR\ from previous time slices, and the accumulation of \SNR, in total the
\lloid\ algorithm requires
%
\begin{multline}
\label{eq:lloid-flops}
\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s \\ + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
\end{multline}
%
\flops.  The second sum is carried out over the set of distinct sample rates
(except for the base sample rate) rather than over the time slices themselves,
as we have found that it is sometimes desirable to place multiple adjacent time
slices at the same sample rate in order to keep the size of matrices that enter
the SVD manageable.  Here we have assumed that the
decimation filters are connected in parallel, converting from the base sample
rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and
that the interpolation filters are connected in cascade fashion with each
interpolation filter stepping from the sample rate of one time slice to the
next.

@@

We examine the performance of the \lloid\ algorithm on a small region of
compact binary parameter space centered on typical \NS--\NS\
masses.  We begin by constructing a template bank that spans component masses
from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO\ noise
power spectrum~\citep{ALIGONoise}\footnote{\url{http://dcc.ligo.org/cgi-bin/DocDB/ShowDocument?docid=T0900288&version=3}}.  Waveforms are generated in the frequency domain in the stationary phase approximation at (post)$^{3.5}$-Newtonian order in phase and Newtonian order in amplitude \citep[the TaylorF2 waveforms described in][]{TaylorF2}.  Templates are truncated at 10~Hz, where the projected
sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
This results in a grid of 98,544 points, or
$2 \times 98,544 = 197,088$~templates.  Then we create sub-banks by partitioning
the parameter space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this
procedure. We concentrate on a sub-bank with 657 points with chirp masses
between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
\begin{figure}[h]
	\plotone{figures/tmpltbank}
	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
case study, consisting of component masses $m_1$ and $m_2$, between 1 and 3~$M_\odot$, and
chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
\end{figure}
With this sub-bank we are able to construct an efficient time slice decomposition
that consists of 11 time slices with sample rates between 32 and 4096 Hz summarized
in Table~\ref{tab:time_slices}.
\begin{table*}
\caption{\label{tab:time_slices} Filter Design Sub-Bank of 1314 Templates.}
\begin{center}
\begin{tabular}{crr@{,\,}lc*{6}{r}}
\tableline\tableline
\\ [-2ex]
& $f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
\cline{6-11}
\\[-2.5ex]
& (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline
\multirow{11}{*}{\includegraphics{figures/envelope}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[0.5em]
& 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[0.5em]
& 256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\[0.5em]
& 128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\[0.5em]
& 64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\[0.5em]
& 64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\[0.5em]
& 64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\[0.5em]
& 32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\[0.5em]
& 32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\[0.5em]
& 32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\[0.5em]
& 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\[0.5em]
\tableline
\end{tabular}
\tablecomments{From
left to right, this table shows the sample rate, time interval, number of
samples, and number of orthogonal templates for each time slice.  We vary \SVD\
tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
\end{center}
\end{table*}
We use this sub-bank and decomposition for the remainder of this section.

>>>>

From the combination of the time slice decomposition in
Equation~(\ref{eq:time-sliced-templates}) and the \SVD\ defined in
Equation~(\ref{eq:svddecomp}), we define the early-warning output accumulated
up to time slice $s$ using the recurrence relation,
%
% orthogonal decomposition filtering
%
\begin{equation}
	\rho_i^s [k] =%
		% Interpolation SNR
		\overbrace{
			\left(H^\uparrow \rho_i^{s+1}\right)[k]
		}^\textrm{\clap{S/N from previous time slices}}
		% Plus ...
		+
		% Reconstruction
		\underbrace{
			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
		}_\textrm{\clap{reconstruction}}
		% Orthogonal FIR filter
		\overbrace{
			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
\end{equation}
%
%
Observe that the early-warning output for time slice 0, $\rho_i^0[k]$,
approximates the \SNR\ of the original templates.
%
%
\begin{figure*}[h!]
	\begin{center}
		\includegraphics{figures/lloid-diagram}
		\caption{\label{fig:pipeline} Schematic of \lloid\ pipeline illustrating
signal flow.  Circles with arrows represent interpolation
\protect\includegraphics{figures/upsample-symbol} or decimation
\protect\includegraphics{figures/downsample-symbol}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol}.  Squares
\protect\includegraphics{figures/fir-symbol} stand for \fir\ filters.  Sample
rate decreases from the top of the diagram to the bottom.  In this diagram, each
time slice contains three \fir\ filters that are linearly combined to produce
four output channels.  In a typical pipeline, the number of \fir\ filters is
much less than the number of output channels.}
	\end{center}
\end{figure*}
%
%
The signal flow diagram in Figure~\ref{fig:pipeline} illustrates this
recursion relation as a multirate filter network with a number of early-warning outputs.

@@

Taking into account the decimation of the detector data, the orthogonal \fir\
filters, the reconstruction of the time-sliced templates, the interpolation of
\SNR\ from previous time slices, and the accumulation of \SNR, in total the
\lloid\ algorithm requires
%
\begin{equation}
\label{eq:lloid-flops}
\sum_{\mathclap{s=0}}^{\mathclap{S-1}} \left( 2 \slicessamps \numsvdtmps + 2 \numtmps \numsvdtmps + \numtmps \right) f^s + 2\sum_{\mathclap{f^s \in \{f^k \, : \, 0 < k < S\}}} \left( N^\shortdownarrow f^0 + \numtmps N^\shortuparrow f^{s-1}\right)
\end{equation}
%
\flops.  The second sum is carried out over the set of distinct sample rates
(except for the base sample rate) rather than over the time slices themselves,
as we have found that it is sometimes desirable to place multiple adjacent time
slices at the same sample rate in order to keep the size of matrices that enter
the SVD manageable.  Here we have assumed that the
decimation filters are connected in parallel, converting from the base sample
rate $f^0$ to each of the time slice sample rates $f^1$, $f^2$, $\dots$, and
that the interpolation filters are connected in cascade fashion with each
interpolation filter stepping from the sample rate of one time slice to the
next.

@@

We examine the performance of the \lloid\ algorithm on a small region of
compact binary parameter space centered on typical \NS--\NS\
masses.  We begin by constructing a template bank that spans component masses
from 1~to~3~$M_\odot$ using a simulated Advanced \LIGO\ noise
power spectrum~\citep{ALIGONoise}\footnote{\url{http://dcc.ligo.org/cgi-bin/DocDB/ShowDocument?docid=T0900288&version=3}}.  Waveforms are generated in the frequency domain in the stationary phase approximation at (post)$^{3.5}$-Newtonian order in phase and Newtonian order in amplitude \citep[the TaylorF2 waveforms described in][]{TaylorF2}.  Templates are truncated at 10~Hz, where the projected
sensitivity of Advanced \LIGO\ is interrupted by the ``seismic wall.''
This results in a grid of 98,544 points, or
$2 \times 98,544 = 197,088$~templates.  Then we create sub-banks by partitioning
the parameter space by chirp mass.  Figure \ref{fig:tmpltbank} illustrates this
procedure. We concentrate on a sub-bank with 657 points with chirp masses
between 1.1955 and 1.2045~$M_\odot$, or $2 \times 657 = 1314$~templates.
\begin{figure}[h]
	\plotone{figures/tmpltbank}
	\caption{\label{fig:tmpltbank}Source parameters selected for sub-bank used in this
case study, consisting of component masses $m_1$ and $m_2$, between 1 and 3~$M_\odot$, and
chirp masses $\mathcal{M}$ between 1.1955 and 1.2045~$M_\odot$.}
\end{figure}
With this sub-bank we are able to construct an efficient time slice decomposition
that consists of 11 time slices with sample rates between 32 and 4096 Hz summarized
in Table~\ref{tab:time_slices}.
\begin{table*}
\caption{\label{tab:time_slices} Filter Design Sub-Bank of 1314 Templates.}
\begin{center}
\begin{tabular}{crr@{,\,}lc*{6}{r}}
\tableline\tableline
\\ [-1.5ex]
& $f^s$ & $[t^s$&$t^{s+1})$ & &\multicolumn{6}{c}{$-\log_{10}$ (1$-$\SVD\ tolerance)} \\% [1ex]
\cline{6-11}
\\[-2ex]
& (Hz) & \multicolumn{2}{c}{(s)} & $N^s$ & $1$ & $2$ & $3$ & $4$ & $5$ & $6$ \\ \tableline\\
\multirow{11}{*}{\includegraphics{figures/envelope}} & 4096 & [0&0.5) & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\[1em]
& 512 & [0.5&4.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\[1em]
& 256 & [4.5&12.5) & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\[1em]
& 128 & [12.5&76.5) & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\[1em]
& 64 & [76.5&140.5) & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\[1em]
& 64 & [140.5&268.5) & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\[1em]
& 64 & [268.5&396.5) & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\[1em]
& 32 & [396.5&460.5) & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\[1em]
& 32 & [460.5&588.5) & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\[1em]
& 32 & [588.5&844.5) & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\[1em]
& 32 & [844.5&1100.5) & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\[1em]
\tableline
\end{tabular}
\tablecomments{From
left to right, this table shows the sample rate, time interval, number of
samples, and number of orthogonal templates for each time slice.  We vary \SVD\
tolerance from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.}
\end{center}
\end{table*}
We use this sub-bank and decomposition for the remainder of this section.
